Finding shortest and closest vectors in a lattice of Voronoi's first kind

### 2014-10-22 talk/presentation.tex
minor

Longstanding problem in number theory and computer science:
\begin{itemize}
\item integer programming {\tiny (Lenstra~et.~al. 1983, Kannan 1987)}
\item the factoring of polynomials over rationals {\tiny (Lenstra~et.~al. 1982)}
\item checking solvability by radicals {\tiny (Landau and Miller 1985)}
\item solving low density subset sum problems {\tiny (Coster~et.~al. 1992)}
\end{itemize}
}

@@

Communications applications:
\begin{itemize}
\item minimum distance of lattice codes {\tiny (Conway and Sloane 1999, Erez~et.~al. 2004)}
\item capacity analysis of network codes {\tiny (Nazer and Gastpar 2011)}
\end{itemize}
}

@@

\frame{
\frametitle{The shortest vector problem}
\begin{itemize}
\item NP-hard for arbitrary lattices.
\item Easier for specific lattices.
\item Short vectors are easy to find in the \term{root lattices} $\ints^n$, $A_n$, and $D_n$.
\item We will show that the problem is relatively easy to solve for lattices of \term{Voronoi's first kind}.
\end{itemize}
}

@@

\frame{
\frametitle{The closest lattice point problem}
Communications applications:
\begin{itemize}
\item spectrally efficient codes {\tiny (Conway and Sloane 1983, Erez~et.~al.~2004)},
\item vector quantisation and compression {\tiny (Conway~and~Sloane~1982, Erez~et.~al.~2005)},
\item multi-antenna communication (MIMO) {\tiny (El Gamal~et.~al.~2004, Wubben~et.~al.~2011)},
\item equalisation of delay spread channels {\tiny (Zamir~et.~al.~2002, Damen~et.~al.~2003)},
\item network coding {\tiny (Nazer and Gastpar 2011)},
\item blind channel estimation {\tiny (Ryan~et.~al. 2007, McKilliam~2014)}.
\end{itemize}

@@

\frame{
\frametitle{The closest lattice point problem}
\begin{itemize}
\item Applications require fast solutions to the closest lattice point problem.
\item NP-hard for arbitrary lattices.
\item Easier for specific lattices.
\item Fast algorithms exist for the \term{root lattices} $\ints^n$, $A_n$, and $D_n$.
\item We will describe a fast algorithm to compute a closest point in lattices of \term{Voronoi's first kind}.
\end{itemize}
}

@@

\frame{
\frametitle{Lattices of Voronoi's first kind}
\begin{theorem}[Conway and Sloane (1992)]
Let $\Lambda$ be a $n$-dimensional lattice of Voronoi's first kind with obtuse superbase $b_1, \dots, b_{n+1}$.  The relevant vectors in $\Lambda$ are of the form 
\[
\sum_{i \in I} b_i
\] 
where $I$ is a strict subset of $\{1, 2, \dots, n+1\}$ and $I$ is not empty.
\end{theorem}

@@

\begin{itemize}
\item Are any existing cryptographic schemes weakened?
\item Can good codes or quantisers be constructed from lattices of Voronoi's first kind?
\item Can MIMO communications systems benefit?
\item Do applications such as global positioning, frequency estimation, etc., involve lattices of Voronoi's first kind?
\item Are there subfamilies of Voronoi's first kind that admit even faster algorithms?
\item Are there other families of lattices for which similar techniques lead to polynomial time algorithms?
\end{itemize}

>>>>

Longstanding problem in number theory and computer science:
\begin{itemize}
\item sphere packing {\tiny (Voronoi 1908, Conway and Sloane 1999)}
\item integer programming {\tiny (Lenstra~et.~al. 1983, Kannan 1987)}
\item the factoring of polynomials over rationals {\tiny (Lenstra~et.~al. 1982)}
\item checking solvability by radicals {\tiny (Landau and Miller 1985)}
\item solving low density subset sum problems {\tiny (Coster~et.~al. 1992)}
\end{itemize}
}

@@

Communications applications:
\begin{itemize}
\item minimum distance of lattice codes {\tiny (Conway and Sloane 1982, Erez~et.~al. 2004)}
\item capacity analysis of network codes {\tiny (Nazer and Gastpar 2011)}
\end{itemize}
}

@@

\frame{
\frametitle{The shortest vector problem}
\begin{itemize}
\item NP-hard for arbitrary lattices.
\item Easier for specific lattices.
\item Short vectors are easy to find in the \term{root lattices} $\ints^n$, $A_n$, and $D_n$.
\item We will show that the problem is relatively easy to solve for lattices of \term{Voronoi's first kind} {\tiny (Voronoi~1908, Conway~and~Sloane~1992, McKilliam~and~Grant~2012)}.
\end{itemize}
}

@@

\frame{
\frametitle{The closest lattice point problem}
Communications applications:
\begin{itemize}
\item spectrally efficient codes {\tiny (Conway and Sloane 1983, Erez~et.~al.~2004)},
\item quantisation and compression {\tiny (Conway~and~Sloane~1982, Erez~et.~al.~2005)},
\item multi-antenna communication (MIMO) {\tiny (El Gamal~et.~al.~2004, Wubben~et.~al.~2011)},
\item equalisation of delay spread channels {\tiny (Zamir~et.~al.~2002, Damen~et.~al.~2003)},
\item network coding {\tiny (Nazer and Gastpar 2011)},
\item blind channel estimation {\tiny (Ryan~et.~al. 2007, McKilliam~2014)}.
\end{itemize}

@@

\frame{
\frametitle{The closest lattice point problem}
\begin{itemize}
\item Require fast solutions to the closest lattice point problem.
\item NP-hard for arbitrary lattices.
\item Easier for specific lattices.
\item Fast algorithms exist for the \term{root lattices} $\ints^n$, $A_n$, and $D_n$.
\item We will describe a fast algorithm to compute a closest point in lattices of \term{Voronoi's first kind}.
\end{itemize}
}

@@

\frame{
\frametitle{Lattices of Voronoi's first kind}
\begin{theorem}[Conway and Sloane 1992]
Let $\Lambda$ be a $n$-dimensional lattice of Voronoi's first kind with obtuse superbase $b_1, \dots, b_{n+1}$.  The relevant vectors in $\Lambda$ are of the form 
\[
\sum_{i \in I} b_i
\] 
where $I$ is a strict subset of $\{1, 2, \dots, n+1\}$ and $I$ is not empty.
\end{theorem}

@@

\begin{itemize}
\item Are any existing cryptographic schemes weakened?
\item Can good codes or quantisers be constructed from lattices of Voronoi's first kind?
\item Can MIMO communication systems benefit?
\item Do applications such as global positioning, frequency estimation, etc., involve lattices of Voronoi's first kind?
\item Are there subfamilies of Voronoi's first kind that admit even faster algorithms?
\item Are there other families of lattices for which similar techniques lead to polynomial time algorithms?
\end{itemize}
